---
title: "Untitled"
author: "Prakhar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Loading and Preprocessing the Data

```{r load-data}
library(ggplot2)
data <- read.csv("activity.csv")
data$date <- as.Date(data$date)
```

## Total Number of Steps per Day

```{r total-steps}
total_steps <- aggregate(steps ~ date, data, sum)
ggplot(total_steps, aes(x = date, y = steps)) + 
  geom_histogram(stat = "identity", fill = "blue", alpha = 0.7) +
  labs(title = "Total Steps per Day", x = "Date", y = "Steps")
```

```{r mean-median}
mean_steps <- mean(total_steps$steps, na.rm = TRUE)
median_steps <- median(total_steps$steps, na.rm = TRUE)
mean_steps
median_steps
```

## Average Daily Activity Pattern

```{r daily-pattern}
avg_steps <- aggregate(steps ~ interval, data, mean, na.rm = TRUE)
ggplot(avg_steps, aes(x = interval, y = steps)) + 
  geom_line(color = "red", linewidth = 1) +
  labs(title = "Average Daily Activity Pattern", x = "Interval", y = "Average Steps")
```

```{r max-interval}
max_interval <- avg_steps[which.max(avg_steps$steps), ]
max_interval
```

## Imputing Missing Values

```{r missing-values}
missing_values <- sum(is.na(data$steps))
missing_values
imputed_data <- data
imputed_data$steps[is.na(imputed_data$steps)] <- mean(imputed_data$steps, na.rm = TRUE)
```

## Histogram After Imputation

```{r histogram-imputed}
total_steps_imputed <- aggregate(steps ~ date, imputed_data, sum)
ggplot(total_steps_imputed, aes(x = date, y = steps)) + 
  geom_histogram(stat = "identity", fill = "green", alpha = 0.7) +
  labs(title = "Total Steps per Day After Imputation", x = "Date", y = "Steps")
```

## Weekday vs. Weekend Activity Patterns

```{r weekday-weekend}
imputed_data$day_type <- ifelse(weekdays(imputed_data$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
avg_steps_weekday <- aggregate(steps ~ interval + day_type, imputed_data, mean)
ggplot(avg_steps_weekday, aes(x = interval, y = steps, color = day_type)) + 
  geom_line(linewidth = 1) +
  facet_wrap(~day_type, ncol = 1) +
  labs(title = "Weekday vs. Weekend Activity Patterns", x = "Interval", y = "Average Steps")
```

## Conclusion

This analysis examined personal activity monitoring data, showing trends in step counts, identifying the most active intervals, handling missing data, and comparing weekday vs. weekend activity patterns.
